TestEvol: A Tool for Analyzing Test-Suite Evolution

### 2012-11-03 icse2013demo-testevol/paper.tex
Complete pass

\hypersetup{
colorlinks=true,
bookmarksnumbered=true,
pdftitle={Isolating Failure Causes through Test Case Generation},
pdfauthor={Jeremias R\"o\ss{}ler and Gordon Fraser and Andreas Zeller and Alessandro Orso},
pdfsubject={Automated Testing},
pdfkeywords={Automated Debugging, Diagnostics, Testing tools}
}

@@

% \note{This is a VERY rough draft that I put together by copying and
%   pasting from the FSE paper. I left some specific TODOs in, but the
%   paper really needs a major pass.}
% \todo{Restructure the content to make sure we comply with what is
%   requested from a demo paper: Tool-based demonstrations describe
%   novel aspects of early prototypes or mature tools. The tool
%   demonstrations must communicate clearly the following information to
%   the audience:\\
%   - the envisioned users;\\
%   - the software engineering challenge it proposes to address;\\
%   - the methodology it implies for its users; and\\
%   - the results of validation studies already conducted for mature tools, or the design of planned studies for early prototypes.}
% \todo{Reduce the description of the approach and extend the
%   description of the implementation. Leandro, can you add some details
%   there?}
% \todo{Add snapshots.}

\begin{abstract}
  Test suites, once created, rarely remain static. Just like the
  application they are testing, they evolve throughout their lifetime.
  Test obsolescence is probably the most known reason for test-suite
  evolution---test cases cease to work because of changes in the code
  and must be suitably repaired. There are several reasons why it is
  important to achieve a thorough understanding of how test cases
  evolve in practice. In particular, researchers who develop automated
  test-repair techniques---an increasingly active research area---can
  use such understanding to develop more effective repair techniques
  that are applicable in real-world scenarios. More generally,
  analyzing test-suite evolution can help testers better understand
  how test cases are modified during maintenance and improve the test
  evolution process, an extremely time consuming activity for any non
  trivial test suite.  Unfortunately, there are no existing tools that
  facilitate investigation of test-suite evolution. To tackle this
  problem, we developed \tool, a tool that enables the systematic
  study of test-suite evolution for Java programs and JUnit test
  cases.  This demonstration presents \tool and demonstrates its
  usefulness and practical applicability by showing how \tool\ can be
  used on real-world software and unit test suites.

@@

Test suites are not static entities: they constantly evolve along with
the application they test. For instance, new tests can be added to
test new functionality and existing tests can be refactored, repaired,
or deleted.  In particular, changes in the application can break test
cases---in some cases, even a small change in the application code can
affect a large number of tests.

A broken test, if it covers a valid functionality, should ideally be
repaired. Alternatively, if the repair is unduly complex to perform or
if the test was designed to cover a functionality that no longer
exists in the application, the test should be removed from the test
suite. To illustrate with an example,
Figure~\ref{fig:pmd-javatokenizertest} shows two versions of a unit
test case from the test suite of \subject{PMD}, one of the programs we
analyzed in our previous work~\cite{pinto12}. A change in
\subject{PMD}'s API broke the original version of the test case, which
had to be fixed by adding a call to method
\lang{SourceCode.readSource} and removing one parameter from the call
to method \lang{Tokenizer.tokenize} (lines 6 and 7 in
Figure~\ref{fig:pmd-javatokenizertest}(b), respectively).

@@

We believe that, to develop effective techniques for assisting manual
test repair, we must first understand how test suites evolve in
practice. That is, we must understand when and how tests are created,
removed, and modified. This is a necessary, preliminary step because
it can (1) provide evidence that test cases do get repaired, (2)
support the hypothesis that test repairs can be (at least partially)
automated, and (3) suitably direct research efforts.  Without such
understanding, we risk to develop techniques that may not be generally
applicable and may not perform the kind of repairs that are actually
needed in real-world software systems.

More generally, the ability to analyze how test suites evolve can be
beneficial for developers, as it can help them better understand the
cost and tradeoffs of test evolution. Similarly, project managers can
use information about test-suite evolution to get insight into the
test maintenance process (\eg how often tests are deleted and cause
oss of coverage, how often tests must be adapted because of changes in
the system's API) and ultimately improve it.

Until recently there were no empirical studies in the literature that
investigated how unit test suites evolve and tools that could support
such studies. To address this issue, we defined a technique that
combines various static- and dynamic-analysis techniques to compute
the differences between the test suites associated with two versions
of a program and categorize such changes along two dimensions: the
static differences between the tests in the two test suites and the
behavioral differences between such tests~\cite{pinto12}. Applying our
technique to a number of real-world programs, we were able to discover
several important aspects of test evolution. For example, we found
evidence that, although test repairs are a relatively small fraction
of the activities performed during test evolution, they are indeed
relevant.  We also found that repair techniques that just focus on
oracles (\ie assertions) are likely to be inapplicable in many cases,
that test cases are rarely removed because they are difficult to fix,
but rather because they have become obsolete, and that test cases are
not only added to check bug fixes and test new functionality, as
expected, but also to validate changes in the code.

This demonstration presents a tool that implements our technique and
that we developed to allow other researchers and practitioners to
perform studies such as we the ones we just summarized. The tool,
which is called \tool, enables a systematic study of test-suite
evolution.  Given two versions of a program and corresponding test
suites, \tool\ automatically computes differences in the behavior of
the test suites on the two program versions, classifies the actual
repairs performed between the two versions, and computes the coverage
attained by the tests on the versions.

After presenting \tool's features and technical details, the
demonstration shows how the tool can be used to study how test suites
evolved over the years for a software project.  To do so, we show
examples of application of \tool\ to a set of real-world open-source
software systems.  Specifically, we show how to use \tool\ to
investigate relevant questions on test evolution, such as the
following ones: What types of test-suite changes occur in practice and
with what frequency?  How often do test repairs require complex
modifications of the tests?  Why are tests deleted and added? Overall,
we show how \tool\ is a useful and practically applicable tool for
researchers in the area of test repair (and test evolution in
general), developers and testers who want to better understand their
software maintenance process.

@@

In this section we firs summarize our technique, implemented in the
\tool\ tool, and then discuss the main characteristics of the tool
implementation. A detailed description of the approach can be found
in~\cite{pinto12}.

@@

\begin{itemize}\denseitems

\item $\pass$: The execution of $P$ against $t$ succeeds.

\item $\failce$: The execution of $P$ against $t$ fails because a
  class or method accessed in $t$ does not exist in
  $P$.\footnote{\small These failures can obviously be detected at
    compile-time. For the sake of consistency in the discussion,
    however, we consider such cases to be detected at runtime via
    ``class not found'' or ``no such method'' exceptions. In fact, our
    \tool{} tool detects such failures at runtime by executing the
    tests compiled using the previous version of $P$ on $P$.}

\item $\failre$: The execution of $P$ against $t$ fails due to an
  uncaught runtime exception (\eg a ``null pointer'' exception).

\item $\failae$: The execution of $P$ against $t$ fails due to an
  assertion violation.

\end{itemize}

We use the generic term $\fail$ to refer to failures for which the
distinction among the above three types of failures is unnecessary.

$\covfunc{P}{t}$ is a function that instruments program $P$, executes
test case $t$ on $P$, and returns the set of all statements in $P$
covered by $t$. $\covfunc{P}{T}$ returns the cumulative coverage
achieved on $P$ by all the tests in test suite $T$.

@@

Figure~\ref{fig:study-design}(a) illustrates the scenario in which $t$
is present in the test suites for both the old and the new versions of
the system. To study different cases, we consider whether $t$ is
modified (to $t'$) and, if so, whether the behaviors of $t$ and $t'$
differ. (We have not considered the cases in which $t$ is not modified
because they are irrelevant for studying test evolution.)  For
\textit{behavioral differences}, there are two cases, shown in the two
rows of the table: either $t$ fails on $P'$ and $t'$ passes on $P'$ or
both $t$ and $t'$ pass on $P'$.

\begin{figure}[t]
\centering
\footnotesize
\tabcolsep=3pt
\textbf{(a) Test $t$ exists in $S$ and $S'$ and is modified}
\\ [2pt]
\begin{tabular}{|l||l|}
\hline
%\multicolumn{1}{|c||}{} & \multicolumn{1}{|c}{} &
%\multicolumn{1}{|c|}{$t$ is not} \\
%\multicolumn{1}{|c||}{} & \multicolumn{1}{|c|}{$t$ is modified}\\
%\hline \hline
$\testfunc{P'}{t} = \fail \; \wedge$ &
$t$ is repaired \\
$\testfunc{P'}{t'} = \pass$ &
\multicolumn{1}{r|}{\scriptsize [\catrep{}]}\\
\hline
\multirow{3}{28mm}{$\testfunc{P'}{t} = \pass \; \wedge \testfunc{P'}{t'} = \pass$} &
$t$ is refactored, updated to test a different \\
& scenario, or is made more/less discriminating \\
& \multicolumn{1}{r|}{\scriptsize [\catref{}]} \\
\hline
\end{tabular}
\\ [8pt]
%
\textbf{(b) Test $t$ is removed in $S'$}
\\ [2pt]
\begin{tabular}{|l||l|}
\hline
\multirow{2}{*}{$\testfunc{P'}{t} = \failre \, | \, \failae$} &
$t$ is too difficult to fix \\
& \multicolumn{1}{r|}{\scriptsize [\catdelaere{}]} \\
\hline
\multirow{2}{*}{$\testfunc{P'}{t} = \failce$} & $t$ is obsolete or is
too difficult to fix \\
& \multicolumn{1}{r|}{\scriptsize [\catdelce{}]} \\
\hline
\multirow{2}{*}{$\testfunc{P'}{t} = \pass$} & $t$ is redundant \\
& \multicolumn{1}{r|}{\scriptsize [\catdelp{}]} \\
\hline 
\end{tabular}
\\ [8pt]
%
\textbf{(c) Test $t'$ is added in $S'$}
\\ [2pt]
\begin{tabular}{|l||l|}
  \hline
  \multirow{2}{*}{$\testfunc{P}{t'} = \failre \, | \, \failae$} &
  $t'$ is added to validate a bug fix \\
  & \multicolumn{1}{r|}{\scriptsize [\cataddaere{}]} \\
  \hline
  \multirow{3}{*}{$\testfunc{P}{t'} = \failce$} &
  $t'$ is added to test a new functionality \\
  & or a code refactoring \\
  & \multicolumn{1}{r|}{\scriptsize [\cataddce{}]} \\
  \hline
  \multirow{3}{*}{$\testfunc{P}{t'} = \pass$} &
  %% TODO: Check: $t'$ is added to test a new functionality \\
  $t'$ is added to test an existing feature\\
  & or for coverage-based augmentation \\
  & \multicolumn{1}{r|}{\scriptsize [\cataddp{}]} \\
  \hline
\end{tabular}
\caption{Scenarios considered in our investigation. Given two system
  versions $S = (P, T)$ and $S' = (P', T')$, the three scenarios are:
  (a) $t$ exists in $T$ and $T'$ and is modified, (b) $t$ exists in
  $T$ but not in $T'$, (c) $t'$ exists in $T'$ but not in $T$.}
\vspace*{-5pt}
\label{fig:study-design}
\end{figure}

\paragraph{Category \catrep{} (Repaired Tests)}
\label{sec:category-rep}

The \catrep{} category corresponds to the case where $t$ is repaired
so that, after the modifications, it passes on $P'.$ As discussed in
the Introduction, Figure~\ref{fig:pmd-javatokenizertest} shows an
example of such a test repair.

@@

\paragraph{Category \catref{} (Refactored Tests)}
\label{sec:category-c2}

The \catref{} category captures scenarios in which a test $t$ is
modified in $S'$ even though $t$ passes on $P'$.
%% Listings~\ref{c2-old} and~\ref{c2-new} show an example of one such
%% change from \subject{Commons Math}. Unit test \lang{testDistance} was
%% refactored to invoke method \lang{sqrt()} in class \lang{FastMath}, a
%% newly added class in the new release, instead of the same method in
%% \lang{java.lang.Math}.

@@

\paragraph{Category \catdelaere{} (Hard-To-Fix Tests)}
\label{sec:category-delaere}

%%COMMENT PLEASE READ ME: I guess here we should focus more
%%in the cases that tests are removed because they have become
%%obsolete. Actually, in the examples I was not able to find one
%%case in which it was removed because it was too difficult to
%% be fixed. It was always obsolete tests.

This category includes tests that fail on $P'$ with a runtime
exception or an assertion violation. These may be instances where the
tests should have been fixed, as the functionality that they test in
$P$ still exists in $P'$, but the tests were discarded instead. One
plausible hypothesis is that tests in this category involve repairs of
undue complexity, for which the investigation of new repair techniques
to aid the developer might be particularly useful.
%% We performed a preliminary investigation of this hypothesis by
%% manually examining ten randomly selected tests in this category and
%% found that all the examined tests were in fact obsolete.

\paragraph{Category \catdelce (Obsolete Tests)}
\label{sec:category-delce}

A test that fails with a compilation error on the new program version
is obsolete because of API changes.
%
%% Listing~\ref{c4-example} illustrates a test in this category taken
%% from \subject{JodaTime}.  This test was deleted because the tested
%% method \lang{Chronology.getBuddhist()} was removed in the subsequent
%% version of \subject{JodaTime}.
%
%% % \vspace{-4pt}
%% \lstinputlisting[language=java,label=c4-example,caption=\small{Unit
%%   test from class TestChronology} (JodaTime v2.0),frame=tb]
%% {code/TestDel-Ce.java}
%% %This method was
%% %already marked as deprecated in the previous version.
%
Although for this category of deletion too, one could postulate that
the tests were removed because they were too difficult to fix, we
believe this not to be the case in most practical occurrences.
Instead, the more likely explanation is that the tests were removed
simply because the tested methods were no longer present.
%% Also in this case, we investigated our hypothesis by manually
%% examining ten randomly selected cases. Indeed, our manual
%% investigation confirmed that, for the cases we analyzed, the tested
%% functionality was either removed or provided through alternative
%% methods, which requires the development of new tests rather than fixes
%% to the existing ones.

\paragraph{Category \catdelp: (Redundant Tests)}
\label{sec:category-delp}

This category includes tests that are removed even though they pass on
$P'$. These tests would typically be redundant, by some measure of
redundancy (\eg code-coverage based). In some cases, a test may be
replaced with a different one.

%% Listing~\ref{c5-example} illustrates an example of one such test,
%% taken from \subject{Commons Lang}. In the new version, this test was
%% replaced by a more sophisticated one, shown in
%% Listing~\ref{c5.2-example}.

%% \begin{figure}[h]
%% \lstinputlisting[language=java,label=c5-example,caption=\small{Unit
%%   test from class MatrixIndexExceptionTest} (Commons Lang
%% v2.1),frame=tb] {code/TestDel-P.java}
%% \vspace{4pt}
%% \end{figure}

%% \lstinputlisting[language=java,label=c5.2-example,caption=\small{Unit
%%   test implemented to replace the removed one} (Commons Lang
%% v2.2),frame=tb] {code/TestDel-PII.java}

@@

\paragraph{Category \cataddaere: (Bug-Fix Tests)}
\label{sec:category-addaere}

This category includes added tests that fail on $P$ with a runtime
exception or an assertion violation. In this case, the functionality
that the added test $t'$ was designed to test exists in $P$ but is not
working as expected (most likely because of a fault). The program
modifications between $P$ and $P'$ would ostensibly have been made to
fix the fault, which causes $t'$ to pass on $P'$. Thus, $t'$ is added
to the test suite to validate the bug fix.

@@

\paragraph{Category \cataddce (New-Features Tests)}
\label{sec:category-addce}

The added tests in this category fail on $P$ with a compilation error,
which indicates that the API accessed by the tests does not exist in
$P$. Thus, the added test $t'$ is created to test new code in $P'$,
where the code could have been added as part of a refactoring or, more
likely, to add new functionality.

%% Listing~\ref{c7-example} illustrates a test from \subject{JFreeChart}
%% that covers a new functionality: method \lang{getMinY()} did not exist
%% in previous versions of class \lang{TimeSeries}.

%% \lstinputlisting[language=java,label=c7-example,caption=\small{Unit
%%   test from class TimeSeriesTests} (JFreeChart v1.0.14),
%% captionpos=t,frame=tb]{code/TestAdd-Ce.java}

\paragraph{Category \cataddp (Coverage-Augmentation Tests)}
\label{sec:category-addp}

This category considers cases where the added test $t'$ passes on
$P$. Clearly, $t'$ would have been a valid test in the old system as
well. One would expect that the addition of $t'$ increases program
coverage (\ie $\covfunc{P'}{T'} \supset \covfunc{P'}{T' -
  \{t'\}}$). Moreover, if $t'$ covers different statements in $P$ and
$P'$ (assuming that there is a way of matching statements between $P$
and $P'$), the plausible explanation is that $t'$ was added to test
the changes made between $P$ and $P'$. However, if $t'$ covers the
same statements in both program versions, it would have been added
purely to increase code coverage (and not to test any added or
modified code).

%% Listing~\ref{c8-example} illustrates an added test case that increases
%% code coverage: the new test method \lang{testFindRangeBounds} covers
%% the case where the parameter of method \lang{findRangeBounds()} is
%% null.

%% \lstinputlisting[language=java,label=c8-example,caption=\small{Unit test
%% from class XYErrorRendererTests} (JFreeChart v1.0.14),frame=tb]
%% {code/TestAdd-P.java}

@@

In our demonstration, we show an implementation of our technique,
\tool, that works on Java programs and JUnit test suites ({\small
  \url{http://www.junit.org/}}). We chose Java and JUnit because the
former is a widely used language, and the latter is the de-facto
standard unit-testing framework for Java.  \tool\ analyzes a sequence
of versions of a software system, where each version can be an actual
release or an internal build and consists of application code and test
code.

\tool\ is implemented as a Java based web application that runs on
Apache Tomcat ({\small \url{http://tomcat.apache.org/}}). Although the
tool can be installed locally, users who want to get familiar with
\tool\ without having to install it can also access it remotely at
{\small \url{http://cheetah.cc.gt.atl.ga.us:8081/testevol/}},
specifying user ``icse'' and password ``icse2013''.

Due to space limitations, we can only provide a few examples of
\tool's graphical interface. (The companion demo video illustrates the
tool interface in full detail.) Figure~\ref{fig:summary} shows the
summary report generated by \tool\ for two pairs of versions of
project \subject{google-gson}. The report shows the number of tests,
both total and per version, that fall in each of the eight categories
defined in our approach.  By clicking on a pair of versions, the user
can obtain detailed report for a specific pair of versions. To
illustrate, Figure~\ref{fig:versiondetails} shows the detailed report
for versions $v.1.1$--$v.1.2$ and a specific category of differences
(\catrep). By further clicking on a test case in the list, users can
also inspect, in the case of modified tests, the actual differences
between the two versions of that test, as shown in
Figure~\ref{fig:summary} for test case \subject{TypeInfoTest}.

@@

The test outcomes, collected by the test-execution engine, and the
test-suite changes, computed by the differencing component, are then
passed to the \emph{test classifier}, which analyzes the information
about test outcomes and test updates to classify each update into one
of the eight test-evolution categories presented in
Section~\ref{sec:test-evolution}. For each pair consisting of a broken
test case and its repaired version, the test classifier also compares
the test cases to identify the types of repair changes---different
types of method-sequence changes and assertion changes---as discussed
in Section~\ref{sec:category-rep}; this analysis is also implemented
using \textsc{wala}.

\tool\ performs a further step for the test cases in categories
\catdelp{}and \cataddp{}. For these tests, the test classifier
leverages the \emph{coverage analyzer} to compute the branch coverage
achieved by each test; this facilitates the investigation of whether
the deleted or added tests cause any variations in the coverage
attained by the new test suite.

@@

\bibliographystyle{abbrv}
{\footnotesize
\bibliography{paper}
}

>>>>

\hypersetup{
colorlinks=true,
bookmarksnumbered=true,
pdftitle={TestEvol: A Tool for Analyzing Test-Suite Evolution},
pdfauthor={Leandro Sales Pinto and Saurabh Sinha and Alessandro Orso},
pdfsubject={Test Evolution},
pdfkeywords={Test Repair, Test-Suite Evolution, Test Maintenance}
}

@@

% Tool-based demonstrations describe novel aspects of early prototypes
% or mature tools. The tool demonstrations must communicate clearly
% the following information to the audience:\\
% - the envisioned users;\\
% - the software engineering challenge it proposes to address;\\
% - the methodology it implies for its users; and\\
% - the results of validation studies already conducted for mature
% tools, or the design of planned studies for early prototypes.}

\begin{abstract}
  Test suites, just like the applications they are testing, evolve
  throughout their lifetime.  One of the main reasons for test-suite
  evolution is test obsolescence: test cases cease to work because of
  changes in the code and must be suitably repaired. There are several
  reasons why it is important to achieve a thorough understanding of
  how test cases evolve in practice. In particular, researchers who
  develop automated test-repair techniques---an increasingly active
  research area---can use such understanding to develop more effective
  repair techniques that are applicable in real-world scenarios. More
  generally, analyzing test-suite evolution can help testers better
  understand how test cases are modified during maintenance and
  improve the test evolution process, an extremely time consuming
  activity for any non-trivial test suite.  Unfortunately, there are
  no existing tools that facilitate investigation of test evolution.
  To tackle this problem, we developed \tool, a tool that enables the
  systematic study of test-suite evolution for Java programs and JUnit
  test cases. This demonstration presents \tool\ and illustrates its
  usefulness and practical applicability by showing how \tool\ can be
  successfully used on real-world software and test suites.

@@

Test suites are not static entities: they constantly evolve along with
the applications they test. For instance, new tests can be added to
test new functionality, and existing tests can be refactored,
repaired, or deleted. Often, test cases have to be modified because
changes in the application break them. In these cases, if the broken
test covers a valid functionality, it should ideally be repaired.
Alternatively, if the repair is unduly complex to perform, or if the
test was designed to cover a functionality that no longer exists in
the application, the test should be removed from the test suite. To
illustrate with an example, Figure~\ref{fig:pmd-javatokenizertest}
shows two versions of a unit test case from the test suite of
\subject{PMD}, one of the programs we analyzed in previous
work~\cite{pinto12}.  A change in \subject{PMD}'s API broke the
original version of the test case, which had to be fixed by adding a
call to method \lang{SourceCode.readSource} and removing one parameter
from the call to method \lang{Tokenizer.tokenize} (lines 6 and 7 in
Figure~\ref{fig:pmd-javatokenizertest}(b), respectively).

@@

We believe that, to develop effective techniques for assisting manual
test repair, we must first understand how test suites evolve in
practice.  That is, we must understand when and how tests are created,
removed, and modified. Such understanding can (1) provide evidence
that test cases do get repaired, (2) support the hypothesis that test
repairs can be (at least partially) automated, and (3) suitably guide
research efforts.  Otherwise, we risk to develop techniques that may
not be generally applicable and may not perform the kind of repairs
that are actually needed in real-world software systems.

More generally, the ability to analyze how test suites evolve can be
beneficial for developers, as it can help them better understand the
cost and tradeoffs of test evolution. Similarly, project managers can
use information about test-suite evolution to get insight into the
test maintenance process (\eg how often tests are deleted and cause
loss of coverage or how often tests must be adapted because of changes
in the system's API) and ultimately improve it.

Until recently, there were no empirical studies in the literature that
investigated how unit test suites evolve, and no tools that could
support such studies existed. To address this issue, we defined a
technique that combines various static- and dynamic-analysis
techniques to (1) compute the differences between the test suites
associated with two versions of a program and (2) categorize such
changes along two dimensions: the static differences between the tests
in the two test suites and the behavioral differences between such
tests~\cite{pinto12}. Applying our technique to a number of real-world
programs, we were able to discover several important aspects of test
evolution. For example, we found evidence that, although test repairs
are a relatively small fraction of the activities performed during
test evolution, they are indeed relevant.  We also found that repair
techniques that just focus on oracles (\ie assertions) are likely to
be inapplicable in many cases, that test cases are rarely removed
because they are difficult to fix, but rather because they have become
obsolete, and that test cases are not only added to check bug fixes
and test new functionality, as expected, but also to validate modified
code.

This demonstration presents a tool, \tool, that implements our
technique and that we developed to allow other researchers and
practitioners to perform additional studies on test evolution. Given
two versions of a program and corresponding test suites, \tool\
automatically computes and classifies both static and behavioral
differences between the test suites for the two program versions.
\tool\ can identify deleted, added, and repaired test cases, together
with the effect that such deletions, additions, and repairs have on
code coverage.

After presenting \tool's features and technical details, the
demonstration shows how the tool can be used to study the evolution,
over the years, of a software project's test suite.  To do so, we show
examples of application of \tool\ to several real-world open-source
software systems. In particular, we show how to use \tool\ to
investigate relevant questions on test evolution, such as what types
of test-suite changes occur in practice and with what frequency, how
often test repairs require complex modifications of the tests, and why
tests are deleted and added. We demonstrate how \tool\ is a useful and
practically applicable tool for researchers and practitioners
interested in test repair (and test evolution in general), and for
developers and testers who want to better understand their test
maintenance process.

@@

In this section we first summarize our technique, implemented in the
\tool\ tool, and then discuss the main characteristics of the tool
implementation. A detailed description of the approach can be found
in~\cite{pinto12}.

@@

\noindent
1) $\pass$: The execution of $P$ against $t$ succeeds.

\noindent
2) $\failce$: The execution of $P$ against $t$ fails because a class
or method accessed in $t$ does not exist in $P$.\footnote{\small These
  failures can obviously be detected at compile-time. For consistency
  in the discussion, however, we consider such cases to be detected at
  runtime via ``class not found'' or ``no such method'' exceptions. In
  fact, \tool\ detects such failures at runtime by executing the tests
  compiled using the previous version of $P$ on $P$.}

\noindent
3) $\failre$: The execution of $P$ against $t$ fails due to an
uncaught runtime exception (\eg a ``null pointer'' exception).

\noindent
4) $\failae$: The execution of $P$ against $t$ fails due to an
assertion violation.

We use the generic term $\fail$ to refer to failures for which the
distinction among different types of failures is unnecessary.

% $\covfunc{P}{t}$ is a function that instruments program $P$, executes
% test case $t$ on $P$, and returns the set of all statements in $P$
% covered by $t$. $\covfunc{P}{T}$ returns the cumulative coverage
% achieved on $P$ by all the tests in test suite $T$.

@@

Figure~\ref{fig:study-design}(a) illustrates the scenario in which $t$
is present in the test suites for both the old and the new versions of
the system. To study different cases, we consider whether $t$ is
modified (to $t'$) and, if so, whether the behaviors of $t$ and $t'$
differ. For \textit{behavioral differences} there are two cases, shown
in the two rows of the table: either $t$ fails on $P'$ and $t'$ passes
on $P'$ or both $t$ and $t'$ pass on $P'$.

\begin{figure}[t]
\centering
\footnotesize
\tabcolsep=3pt
\textbf{(a) Test $t$ exists in $S$ and $S'$ and is modified}
\\ [2pt]
\begin{tabular}{|l||l|}
\hline
%\multicolumn{1}{|c||}{} & \multicolumn{1}{|c}{} &
%\multicolumn{1}{|c|}{$t$ is not} \\
%\multicolumn{1}{|c||}{} & \multicolumn{1}{|c|}{$t$ is modified}\\
%\hline \hline
$\testfunc{P'}{t} = \fail \; \wedge$ &
$t$ is repaired \\
$\testfunc{P'}{t'} = \pass$ &
\multicolumn{1}{r|}{\scriptsize [\catrep{}]}\\
\hline
\multirow{3}{28mm}{$\testfunc{P'}{t} = \pass \; \wedge \testfunc{P'}{t'} = \pass$} &
$t$ is refactored, updated to test a different \\
& scenario, or is made more/less discriminating \\
& \multicolumn{1}{r|}{\scriptsize [\catref{}]} \\
\hline
\end{tabular}
\\ [8pt]
%
\textbf{(b) Test $t$ is removed in $S'$}
\\ [2pt]
\begin{tabular}{|l||l|}
\hline
\multirow{2}{*}{$\testfunc{P'}{t} = \failre \, | \, \failae$} &
$t$ is too difficult to fix \\
& \multicolumn{1}{r|}{\scriptsize [\catdelaere{}]} \\
\hline
\multirow{2}{*}{$\testfunc{P'}{t} = \failce$} & $t$ is obsolete or is
too difficult to fix \\
& \multicolumn{1}{r|}{\scriptsize [\catdelce{}]} \\
\hline
\multirow{2}{*}{$\testfunc{P'}{t} = \pass$} & $t$ is redundant \\
& \multicolumn{1}{r|}{\scriptsize [\catdelp{}]} \\
\hline 
\end{tabular}
\\ [8pt]
%
\textbf{(c) Test $t'$ is added in $S'$}
\\ [2pt]
\begin{tabular}{|l||l|}
  \hline
  \multirow{2}{*}{$\testfunc{P}{t'} = \failre \, | \, \failae$} &
  $t'$ is added to validate a bug fix \\
  & \multicolumn{1}{r|}{\scriptsize [\cataddaere{}]} \\
  \hline
  \multirow{3}{*}{$\testfunc{P}{t'} = \failce$} &
  $t'$ is added to test a new functionality \\
  & or a code refactoring \\
  & \multicolumn{1}{r|}{\scriptsize [\cataddce{}]} \\
  \hline
  \multirow{3}{*}{$\testfunc{P}{t'} = \pass$} &
  %% TODO: Check: $t'$ is added to test a new functionality \\
  $t'$ is added to test an existing feature\\
  & or for coverage-based augmentation \\
  & \multicolumn{1}{r|}{\scriptsize [\cataddp{}]} \\
  \hline
\end{tabular}
\caption{Scenarios considered in our approach, given two system
  versions $S = (P, T)$ and $S' = (P', T')$: (a) $t$ exists in $T$ and
  $T'$ and is modified, (b) $t$ exists in $T$ but not in $T'$, (c)
  $t'$ exists in $T'$ but not in $T$.}
\vspace*{-12pt}
\label{fig:study-design}
\end{figure}

\textit{a) Category \catrep{} (Repaired Tests)} corresponds to cases
where $t$ is repaired so that, after the modifications, it passes on
$P'.$ (See Figure~\ref{fig:pmd-javatokenizertest} and discussion in
Section~\ref{sec:intro}.)

@@

\textit{b) Category \catref{} (Refactored Tests)} captures
scenarios in which a test $t$ is modified in $S'$ even though $t$
passes on $P'$.
%% Listings~\ref{c2-old} and~\ref{c2-new} show an example of one such
%% change from \subject{Commons Math}. Unit test \lang{testDistance} was
%% refactored to invoke method \lang{sqrt()} in class \lang{FastMath}, a
%% newly added class in the new release, instead of the same method in
%% \lang{java.lang.Math}.

@@

\textit{a) Category \catdelaere{} (Hard-To-Fix Tests)} includes tests
that fail on $P'$ with a runtime exception or an assertion
violation. These may be instances where the tests should have been
fixed, as the functionality that they test in $P$ still exists in
$P'$, but the tests were discarded instead. One plausible hypothesis
is that tests in this category involve repairs of undue complexity,
for which the investigation of new repair techniques to aid the
developer might be particularly useful.
%% We performed a preliminary investigation of this hypothesis by
%% manually examining ten randomly selected tests in this category and
%% found that all the examined tests were in fact obsolete.

\textit{b) Category \catdelce (Obsolete Tests)} includes tests that
are obsolete because of API changes, and thus fail with a compilation
error on the new program version.
%
%% Listing~\ref{c4-example} illustrates a test in this category taken
%% from \subject{JodaTime}.  This test was deleted because the tested
%% method \lang{Chronology.getBuddhist()} was removed in the subsequent
%% version of \subject{JodaTime}.
%
%% % \vspace{-4pt}
%% \lstinputlisting[language=java,label=c4-example,caption=\small{Unit
%%   test from class TestChronology} (JodaTime v2.0),frame=tb]
%% {code/TestDel-Ce.java}
%% %This method was
%% %already marked as deprecated in the previous version.
%
Although for this category of deletion too, one could postulate that
the tests were removed because they were too difficult to fix, we
believe this not to be the case in most practical occurrences.
Instead, the more likely explanation is that the tests were removed
simply because the tested methods were no longer present.
%% Also in this case, we investigated our hypothesis by manually
%% examining ten randomly selected cases. Indeed, our manual
%% investigation confirmed that, for the cases we analyzed, the tested
%% functionality was either removed or provided through alternative
%% methods, which requires the development of new tests rather than fixes
%% to the existing ones.

\textit{c) Category \catdelp (Redundant Tests)} includes tests that
are removed even though they pass on $P'$. These tests would typically
be redundant according to some criterion, such as code coverage.

@@

\textit{a) Category \cataddaere (Bug-Fix Tests)} includes added tests
that fail on $P$ with a runtime exception or an assertion
violation. In this case, the functionality that the added test $t'$
was designed to test exists in $P$ but is not working as expected
(most likely because of a fault). The program modifications between
$P$ and $P'$ would ostensibly have been made to fix the fault, which
causes $t'$ to pass on $P'$. Thus, $t'$ is added to the test suite to
validate the bug fix.

@@

\textit{b) Category \cataddce (New-Features Tests)} includes tests
that fail on $P$ with a compilation error, which indicates that the
API accessed by the tests does not exist in $P$. Thus, the added test
$t'$ is created to test new code in $P'$.

\textit{c) Category \cataddp (Coverage-Augmentation Tests)} considers
cases where the added test $t'$ passes on $P$. Clearly, $t'$ would
have been a valid test in the old system as well. One would expect
that the addition of $t'$ increases program coverage. Moreover, if
$t'$ covers different statements in $P$ and $P'$, the plausible
explanation is that $t'$ was added to test the changes made between
$P$ and $P'$. However, if $t'$ covers the same statements in both
program versions, it would have been added purely to increase code
coverage, and not to test any added or modified code.

@@

\tool works on Java programs and JUnit test suites ({\small
  \url{http://www.junit.org/}}). We chose Java and JUnit because the
former is a widely used language, and the latter is the de-facto
standard unit-testing framework for Java.  \tool\ analyzes a sequence
of versions of a software system, where each version can be an actual
release or an internal build and consists of application code and test
code.

\tool\ is implemented as a Java based web application that runs on
Apache Tomcat ({\small \url{http://tomcat.apache.org/}}). Although the
tool can be installed locally, users can also run it without any
install by accessing it remotely at {\small
  \url{http://cheetah.cc.gt.atl.ga.us:8081/testevol/}} and specifying
user ``icse'' and password ``icse2013'' (or requesting an account).

Due to space limitations, we can only provide a few examples of
\tool's graphical interface. (The companion demo video illustrates the
tool interface in full detail.) Figure~\ref{fig:summary} shows the
summary report generated by \tool\ for two pairs of versions of
project \subject{google-gson}. The report shows the number of tests,
both total and per version, that fall in each of the test-evolution
categories considered in our approach.  By clicking on a pair of
versions, the user can obtain a detailed report for a specific pair of
versions.  To illustrate, Figure~\ref{fig:versiondetails} shows the
detailed report for versions $v.1.1$--$v.1.2$ and for a specific
category of differences (\catrep). By further clicking on a test case
in the list, users can also inspect, if it is a modified test, the
actual differences between the two versions of that test, as shown in
Figure~\ref{fig:testdetails} for test case \subject{TypeInfoTest}.

@@

The test outcomes, collected by the test-execution engine, and the
test-suite changes, computed by the differencing component, are then
passed to the \emph{test classifier}, which analyzes the information
about test outcomes and test updates to classify each update into one
of the eight test-evolution categories presented in
Section~\ref{sec:test-evolution}.  For each pair consisting of a
broken test case and its repaired version, the test classifier also
compares the test cases to identify the types of repair changes (\eg
different types of method-sequence changes and assertion changes), as
discussed in Section~\ref{sec:test-mod}. This analysis is also
implemented using \textsc{wala}.

\tool\ performs a further step for the test cases in categories
\catdelp{}and \cataddp{}. For these tests, the test classifier
leverages the \emph{coverage analyzer} to compute the branch coverage
achieved by each test; this facilitates the investigation of whether
the deleted or added tests cause any variations in the coverage
achieved by the new test suite.

@@

\bibliographystyle{abbrv}
{\footnotesize
% \bibliography{paper}
\begin{thebibliography}{1}
